---
title: "Fast Food Nutrition Analysis"
author: ''
date: '2019-04-10'
slug: fast-food-nutrition-analysis
tags: []
categories: []
---

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggthemes)
library(tidytext)
library(purrr)
library(stats)
library(corrr)
library(RColorBrewer)
```

## R Markdown

```{r, include=FALSE}
food<-read_csv("fastfood_calories.csv")
```


```{r, echo=FALSE}
head(food)
```
What is up with the "salad" column?

```{r, echo=FALSE}

levels(as.factor(food$salad))
#All salad levels are "Other." This column can be removed.
food$salad<-NULL
food$X1<-NULL

#convert restaraunt to factor
food$restaurant<-as.factor(food$restaurant)
levels(food$restaurant)
```

```{r,warning=FALSE}
#select only 
# library(microbenchmark)
# microbenchmark(
#     dplyr::select_if(food, is.numeric),
#     Filter(is.numeric, food),
#     purrr::keep(food,is.numeric)
# )
```

## Subsetting Practice

```{r,warning=FALSE}
##replace NA content with 0
food<-food %>% mutate_if(is.numeric, ~replace(., is.na(.), 0))

#using unlist(lapply(...)) is the same as sapply(...)

ind <- as.logical(unlist(lapply(food,is.numeric)))

ind2 <- as.logical(sapply(food,is.numeric))

#subset only numeric cols with logical indices
food_num<-food[which(ind2)]
food_num%>% head()
```


How many entries for each restaraunt? Use fct_infreq to reorder by appearance
```{r,warning=FALSE}
food %>% ggplot(aes(x=fct_infreq(restaurant),fill=restaurant)) + 
  geom_histogram(stat = "count")+ 
  scale_fill_brewer(palette = "Purples") + 
  theme_dark()+
  xlab( "Restaurant")
# Order restaurants by median number of calories
# cals = aggregate(calories ~ restaurant, median, data=salad)
# cals = cals[order(cals$calories, decreasing = T),]
# salad$restaurant = factor(salad$restaurant, levels=as.character(cals$restaurant))

```


```{r}
#items containing bacon
food[str_detect(food$item, "[Bb]acon|BLT|B.L.T"),] 

```


```{r}
bacon<-food %>% mutate(contains_bacon = str_detect(food$item, "[Bb]acon|BLT|B.L.T")) %>%
  group_by(contains_bacon) %>% select_if(is.numeric) %>% summarise_at(
    c("calories","sat_fat","trans_fat","total_fat","sodium","cholesterol"), list(~mean(.)),na.rm=TRUE)
```



```{r}
bacon %>% gather(stat,content,-contains_bacon) %>% ggplot(aes(x=stat,y=content,fill=contains_bacon))+ geom_col(position = "dodge", stat = "identity")+ facet_wrap(~stat,scales="free")+xlab("")
```
```{r,message=FALSE}
#show items that appear more than once
dup<-food %>% select(restaurant,item) %>% duplicated()
food[which(dup),]

##remove duplicate items
food<-food %>% distinct(restaurant,item,.keep_all = T)

food<-food %>% mutate(item_num = row_number())
food
food_words<-food %>%
  unnest_tokens(word, item) %>%
  anti_join(stop_words) %>% filter(str_detect(word, "^[a-z']+$")) %>%
  distinct()

word_freq<-food_words %>% group_by(word) %>%
  summarize(freq = n()) %>%
  arrange(desc(freq))

word_freq %>% mutate(word = reorder(word, freq)) %>%
  head(25) %>%  ggplot(aes(x = word, y = freq)) + geom_segment(aes(
    x = word,
    xend = word,
    y = 0,
    yend = freq
  ), color = "black") + geom_point(size = 3, color = "purple")+coord_flip()
```


```{r}
(word_cal<-food_words %>% group_by(word) %>% filter(length(word)>2) %>% mutate(cal_fat_perc=cal_fat/calories) %>% 
  summarize(freq = n(),avg_cal=mean(calories),avg_cal_fat=mean(cal_fat), avg_calfat_perc=mean(cal_fat_perc)) %>% arrange(avg_cal))

```


```{r}
word_cal %>% head(15) %>%  ggplot(aes(x=fct_reorder(word,avg_calfat_perc),y=avg_cal,fill=avg_calfat_perc))+ geom_bar(stat = "identity",position="dodge") +scale_fill_gradient()+coord_flip()
```


```{r}
word_cal %>% tail(15) %>%  ggplot(aes(x=fct_reorder(word,avg_calfat_perc),y=avg_cal,fill=avg_calfat_perc))+ geom_bar(stat = "identity",position="dodge") +scale_fill_gradient()+coord_flip()
```
Meats: chicken,bacon,beef,turkey,steak

Filter out instances where item title contains "w/o"
```{r}
food<-food %>% filter(!str_detect(item,"w/o")) %>% mutate(
    meat_type = case_when(
      str_detect(item, "[Bb]acon|BLT|B.L.T") & !str_detect(item, "Bacon.*Chicken|Chicken.*Bacon") ~ "bacon",
      str_detect(item, "Bacon.*Chicken|Chicken.*Bacon") ~ "bacon,chicken",
      str_detect(item, "[Cc]hick") ~ "chicken",
      str_detect(item, "[Bb]eef|[Cc]heeseburger|[Hh]amburger") ~ "beef",
      str_detect(item, "[Tt]urkey") ~ "turkey",
      str_detect(item, "[Ss]teak") ~ "steak",
      TRUE~"other")) 

```
```{r}

```



```{r}
meat_stat<-food %>% 
  group_by(meat_type) %>%
  summarise_at(c("calories","total_carb","total_fat","protein"),~mean(.,na.rm = T)) %>%
  gather(stat,mean_content,-meat_type) %>%
  filter(meat_type!="other")

##create order within facet plot
meat_stat<-meat_stat %>% 
  #group by facet var, catg. var
  group_by(stat,meat_type) %>% 
  arrange(desc(mean_content)) %>% 
  ungroup() %>% 
  mutate(meat_type=factor(paste(meat_type,stat, sep = "__"), levels = rev(paste(meat_type,stat, sep = "__")))) 
  

meat_stat %>% 
  ggplot(aes(meat_type,mean_content,fill=stat)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~stat,scales="free") +
  coord_flip()+ scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +scale_fill_brewer()+theme_dark()

```



```{r}
food %>% filter(meat_type!="other",calories<1500) %>%  ggplot()+geom_jitter(aes(x=total_carb,y=calories,color="total_carb"))+ geom_jitter(aes(x=protein,y=calories,color="protein"))+geom_jitter(aes(x=total_fat,y=calories,color="total_fat"))+facet_wrap(~meat_type)+xlab("")+coord_flip()
```



```{r out.width=c('33%','33%','33%'),fig.show='hold'}
ggplotRegression <- function(dat, xvar, yvar){

  fml <- paste(yvar, "~", xvar)

  fit <- lm(fml, dat)

  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) +
    geom_point() +
    stat_smooth(method = "lm", col = "red") +
    labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5)))+coord_flip()
}
ggplotRegression(food,"total_carb", "calories")
ggplotRegression(food,"protein", "calories")
ggplotRegression(food,"total_fat", "calories")
```

Have numeric only saved as food_num
Correlation with calories
```{r,message=FALSE}

mycolors <- colorRampPalette(brewer.pal(13, "RdYlBu"))(13)

cor_df<-correlate(food_num)

cor_df %>% focus(calories) %>% mutate(rowname=reorder(rowname,calories)) %>% ggplot(aes(x=rowname,y=calories,fill=rowname))+geom_bar(stat = "identity")+coord_flip()+scale_fill_manual(values=mycolors)
```


```{r,message=FALSE}
cor_df %>% focus(cal_fat) %>% mutate(rowname=reorder(rowname,cal_fat)) %>% ggplot(aes(x=rowname,y=cal_fat,fill=rowname))+geom_bar(stat = "identity")+coord_flip()+scale_fill_manual(values=mycolors)
```

Want food high protein, low fat,low protein
Which rest. has best selection of high protein, low fat (As percentage?)
Add category bins
How many menu items at McD are l
```{r}
#fat dist
food %>% ggplot(aes(total_fat))+geom_histogram()
food %>% select(restaurant,item,total_fat) %>% top_n(10) %>% arrange(desc(total_fat))
summary(food$total_fat)
```


```{r}
food %>% ggplot(aes(total_carb))+geom_histogram()
```
Pass new variable name (e.g. "tfat_cat") --> use !!new_var:= within function

```{r}
# Function to create new binned categorical variables with specified cutoffs

new_cat<-function(new_var,old_var,low,medium,df){
  df %>% 
  mutate(!!new_var:= case_when(
    old_var<low ~"low",
    old_var<medium~"medium ",
    TRUE~"high"))
}

#create new binned categorical variables for total fat
food <-new_cat("tfat_cat",food$total_fat,10,30,food) 
food %>% group_by(tfat_cat) %>% count()
```


```{r}
(rest_fat<-food %>% group_by(restaurant,tfat_cat) %>% count() %>% 
##add missing low category for BK, Sonic
complete(restaurant,tfat_cat) %>% unique()%>% mutate(n=replace(n,is.na(n), 0)))
```


```{r}
rest_fat %>% ggplot(aes(x=restaurant,y=n,fill=fct_reorder(tfat_cat,n)))+geom_col(position="fill")
```
Burger King & Sonic have no low fat menu items in the dataset. Chick-fil-A has the smallest percentage of high fat items, while Subway has the highest percentange of low fat items.


```

