---
title: DVC Points Chart Analysis (Part 1)
author: ''
date: '2019-07-14'
slug: dvc-points-chart-analysis-part-1
categories: []
tags: []
---
In this series of posts, I will analyze and compare the 2020 points charts for Bay Lake Tower and the The Beach Club. Html versions of the charts (used here) can be found on [DVC Request](https://www.dvcrequest.com/point-charts.asp)

This post (part 1) will detail the code to extract the points charts from the html table format they are published in. I will use the rvest package as well as unpivotr to convert the data into a tabular dataframe. 
```{r,message=FALSE,warning=FALSE}
library(rvest)
library(unpivotr)
library(htmltools)
library(tidyverse)
library(kableExtra)
```

```{r}
#read in html data
link<-"https://www.dvcrequest.com/point-chart/bay_lake_towers_2020.html"
hlink<- read_html(link)

table<-hlink %>% html_nodes("table") %>% 
  html_table(header=F,fill=T)
table <- table[[1]]

#add the table to a dataframe, 
dict <- as.data.frame(table)
cells <- as_cells(table)
cells<-cells %>% filter(col>2,row>1) %>% select(-data_type)

cells %>% head() %>% kable()
```
Although we now have the data in a usable data frame, it is all one long column. Therefore, we need to spread and restructure the dataframe.

```{r}
cellstemp<-
   cells %>% spread(row,chr) %>% filter(!col %in% c('6','10','14'))
room_view<-cellstemp[,2:3]
#seasons
adventure<-cellstemp[,4:6]
choice<-cellstemp[,7:9]
dream<-cellstemp[,10:12]
magic<-cellstemp[,13:15]
premier<-cellstemp[,16:18]
adventure<-cbind(room_view,adventure)
choice<-cbind(room_view,choice)
dream<-cbind(room_view,dream)
magic<-cbind(room_view,magic)
premier<-cbind(room_view,premier)
dfs<-data.frame(Map(c,adventure,choice,dream,magic,premier))
dfs %>% head() %>% kable()
```
That's much better: now the room type, room view, season, and points each have their own columns. Now we can rename these columns and gather the points to a single column
```{r}
dfs<-dfs %>% rename_at(c(1,2,3,4,5),~c("room","view","season","Sun_Th","Fri_Sat")) %>%
  mutate_at(vars('Sun_Th','Fri_Sat'),as.character) %>%
  mutate_at(vars('Sun_Th','Fri_Sat'),as.numeric)

dfl<-dfs %>% gather("weekday", "points",Sun_Th:Fri_Sat)
dfl$room<-as.character(dfl$room)

dfl %>% head() %>% kable()
```
The room column entries need to be renamed as well. 
```{r}
#change naming of room type
dfl[dfl$room == "STUDIO(Sleeps up to 4)","room"] = "Studio"
dfl[dfl$room == "1-BEDROOM\r\n      VACATION HOME(Sleeps up to 5)","room"] = "1-Bedroom"
dfl[dfl$room == "2-BEDROOM\r\n      VACATION HOME(Sleeps up to 9)","room"] = "2-Bedroom"
dfl[dfl$room == "3-BEDROOM\r\n  GRAND VILLA(Sleeps up to 12)","room"] = "3-Bedroom"

dfl %>% head() %>% kable()
```
Great, now the data can be easily explore and plotted.

```{r}
dfl %>% 
  ggplot(aes(x=fct_reorder(room,points),y=points,fill=fct_reorder(view,points)))+
  geom_boxplot()+
  scale_fill_brewer(palette= "OrRd")+
  labs(x="",fill="View")

```

```{r}
dfl %>%
  ggplot(aes(x=fct_reorder(season,points),y=points,fill=fct_reorder(view,points)))+geom_boxplot()+scale_fill_brewer(palette= "OrRd")
```
From the above boxplots, we can see that room type determines points more than does season (i.e. there is less variability within a room than within a season)

```{r}
dfl %>% filter(room %in% c("Studio", "1-Bedroom")) %>%
  group_by(room, season) %>% summarize(avg_point = mean(points)) %>% 
  ungroup() %>%
  ggplot(aes(x = fct_reorder(season, avg_point),y = avg_point,group = room))+ 
  geom_col(aes(fill = room), position = "dodge") + 
  geom_text(aes(label =round(avg_point, 1)),position = position_dodge(0.9),vjust = 0) +
  scale_fill_manual(values = c("skyblue", "grey"))
```

To prepare for the next post, where I will compare BLT and the Beach Club, I will add a column indicating the resort and save the RDS file. 
The same process documented above was done for the Beach Club point chart. This can be found on my github [here](https://github.com/mikaylaedwards/DVC_analysis/blob/master/beachclub_analys.Rmd)
```{r}
df20<-dfl %>% mutate(resort="BLT")
df20%>% saveRDS(file = "BLT2020.rds")