---
title: Wait Times Shiny App + More Visuals
author: ''
date: '2019-09-21'
slug: wait-times-revisited
categories: []
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
library(fs)
library(tidyverse)
library(lubridate)
library(plotly)
library(timetk)
library(tibbletime)
library(timeDate)
library(sugrrants)
```

## Shiny App

```{r eval=FALSE}
##shiny code 

ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            selectInput("ride","Ride",choices = unique(files$ride),
                        multiple = TRUE,
                        selected =c("flight_of_passage","7_dwarfs_train","slinky_dog")),
            dateInput('date1',
                      label = 'Date input 1: yyyy-mm-dd',value = ymd("2019-07-04")),
            dateInput('date2',
                      label = 'Date input 2: yyyy-mm-dd',value = ymd("2019-07-05"))
        ),
        mainPanel(
            # Replace the `plotOutput()` with the plotly version
            plotlyOutput("plot")
        )
    )
)

# Define the server logic
server <- function(input, output) {
    # Replace the `renderPlot()` with the plotly version
    output$plot <- renderPlotly({
        # Convert the existing ggplot2 to a plotly plot
        ggplotly({
            data <- subset(files,
                           ride %in% input$ride &
                               date %in% c(input$date1, input$date2))
            p <- ggplot(data, aes(x=datetime,y=time,color=ride))+
                geom_line()+
                facet_grid(~date,scales="free_x")+
                ggtitle("Compare Wait Times")+
                scale_color_brewer(palette = "Dark2")
        })
    })
}

shinyApp(ui = ui, server = server)
```

The app below allows for comparison of wait times on two different dates. Rides can be selected using the dropdown.
The default shown below compares July 4th to July 5th. 

<br>

<iframe width="1500" height="900" scrolling="yes" frameborder="no"  src="https://mikayla-edwards19.shinyapps.io/Ride_Wait_Times/"> </iframe> 
It is interesting that flight of passage is significantly higher than seven dwarfs on pretty much any day, except July 4th. This likely speaks to the popularity of Magic Kingdom fireworks on the 4th. In other words, more people were probably in Magic Kingdom for the 4th than Animal Kingdom. 

## Code used to produce

Print files names in directory
```{r}
files <- "wait-times"
files<-dir_ls(files)

files
```
I want to merge the datasets but first need to create a new column in each with the ride name. I will extract the ride name from the csv file name and then create the new column
```{r warning=FALSE}
files<-files %>% 
  map_dfr(read_csv, .id = "source")

files %>% head()
```

```{r}
files<-files %>% 
  mutate(ride=str_replace(source,"wait-times/","")) %>% 
  mutate(ride=str_extract(ride,"[^.]+")) %>% 
  select(-source) %>% 
  mutate(date=mdy(date)) %>% 
  select(ride,everything())

files  %>% 
  arrange(desc(date)) %>% 
  group_by(ride) %>% 
  slice(1)
```

Gather actual vs posted 
```{r}
files<-files %>%
  select(everything(),posted=SPOSTMIN,actual=SACTMIN) %>% 
    gather(key="type",value="time",posted:actual) %>% 
    na.omit() %>% 
  filter(time>=0,time <=400)

head(files)
```
Remove times from 1am-5am, as these are incorrent entries. Looking at only flight of passage and slinky dog.
```{r}
files<-files %>% 
  mutate(
    hour=hour(datetime)) %>% 
  filter(! hour %in% c(1:5))

# create tibble time object
```
















```{r}
fp_ts <- files %>%
  filter(ride == "flight_of_passage" | ride=="slinky_dog") %>% 
  as_tbl_time(datetime)


fp_ts<-fp_ts %>% 
  collapse_by(period="hourly",start_date = min(fp_ts$datetime),side="start") %>% 
  group_by(ride,datetime) %>% 
  summarise(
    avgTime=mean(time,na.rm=T)
  ) %>%
  slice(-n()) 

fp_ts<-fp_ts %>% 
  mutate(date=date(datetime))
```


```{r}
```
## Visualize
Overall comparison. Only one year of data is shown for slinky dog because it opened late June 2018. 
```{r}
p1<-fp_ts %>% 
  ggplot(aes(x=date,y=avgTime)) + 
  geom_line()+
  geom_smooth(method = "loess",aes(color=ride))+
  facet_grid(~ride,scales = "free_x")+
  scale_x_date(date_labels = "%b %y")+
  theme(panel.spacing.x = unit(2,"lines"))+
  theme_minimal()


ggplotly(p1,width = 800)
```

#  Time-based Features ----
Append features to get day of week, 
```{r}
fp_aug<-fp_ts %>%
  timetk:: tk_augment_timeseries_signature()

head(fp_aug)
```


```{r}
summer_week<-fp_aug %>%
  filter(date %within% interval(ymd("2019-07-01"), ymd("2019-08-30"))) %>% 
  select(ride, date, avgTime, hour, wday.lbl) %>%
  ggplot(aes(x = hour, y = avgTime,color=wday.lbl)) +
  geom_line(size=.7) +
  facet_grid(rows = vars(ride), cols = vars(wday.lbl))+
  scale_color_manual(values=rainbow(13,s=.8,v=.8,rev=T))+
  theme(legend.position = "none")

ggplotly(summer_week)

```

Sharp drop on Monday night for Flight of Passage this summer. Likely has to do with the fact that extra magic hours were offered at Animal Kingdom on this night of the week starting early July 2019. 

###Frame Calendar! 
```{r}
# Holidays


holidays <- holidayNYSE(year = c(2017, 2018,2019)) %>% ymd()

fp_aug<-fp_aug %>% 
  mutate(isHoliday = case_when(
        date(date) %in% holidays ~ "1",
        TRUE ~ "0"
    ))
```

This summer vs last summer
```{r}
cal_plot<-function(ridename){
  
cal <- fp_aug %>%
  filter(ride==ridename) %>% 
  filter(year %in% c(2018,2019)) %>% 
  filter(month.lbl %in% c("June","July","August")) %>%
  frame_calendar(x = hour, y = avgTime, date = date) %>%
  ggplot(aes(
    x = .hour,
    y = .avgTime,
    group = date,
    colour = isHoliday
  )) +
  geom_line()
  
prettify(cal)  

}


cal_plot("flight_of_passage")
cal_plot("slinky_dog")
```
```{r}
files %>% 
  filter(date==ymd("2019-08-01")) %>% 
  ggplot(aes(x=datetime,y=time,fill=ride))+
  geom_line()
```















