---
title: How to write API packages for R
author: ''
date: '2020-08-17'
draft: true
slug: how-to-write-api-packages-for-r
categories:
  - R package
  - Guide
tags:
  - R
  - Vehicle fleet
  - Govt
  - API
subtitle: ''
summary: ''
authors: []
lastmod: '2020-08-17T17:15:13+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


# Overview

# Installation

`DVLA` is not yet on CRAN. To install, simply run the following in R.

```{r installation, eval=FALSE, include=TRUE}
devtools::install_github("daveyr/DVLA")
```

# Usage
Documentation on functions and examples can be found on its [Github page](https://daveyr.github.io/DVLA/). Here's a more involved example, where we loop through a vector of car registration numbers.


```{r}
# DVLA API in action
# 
# library(DVLA)
#  i <- 1
#  dvla <- NULL
#  for(reg in unique(fleet$reg)){
#    df <- getDVLA(key, reg)
#    dvla <- rbindlist(list(dvla,df), fill = T)
#    i <- i + 1
#    Sys.sleep(0.2) # avoid rate limits
#  }
#  
#  complete <- fleet %>%
# 	left_join(dvla, by = c("reg" = "registrationNumber"))
```

# Backend
So what is happening behind the scenes? How would one go about replicating this type of package for another API?
